---
title: "Fraud Analysis_M-KOP@ @SSESSMENT"
author: "ERICK@"
date: "2024-"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem Statement: Questions to answer:

1)  Which region has the highest Outstanding Loan Balance exposure. Provide reasons and evidence of the analysis. Explain in approximately no less than (80 words)
2)  Giving reasons and rank the region that is highly affected by fraud.Show evidence and explain in approximately (70 words)
3)  Using the data determine the most affected phone model, and in which region. Show evidence and explain in approximately (70 words).
4)  Using the data show the most affected month by fraud. Show evidence and explain in approximately (50 words).
5)  How could we potentially improve the fraud identification process? explain in approximately (100 words).
6)  What operational improvements should we investigate to improve the fraud investigation process? explain in approximately (100 words).
7)  Write an SQL query to replicate the results in Data-Sheet but only getting results for Suwami reg. Use the data on sheet named "Short schema".
8)  Show the process you used to clean the data. Show evidence and explain in approximately (80 words)

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
```

```{r warning=F, message=FALSE}
df <- read_csv("C:/Users/langa/OneDrive/Desktop/Dataset/M-KOPA Fraud Analyst Intern Technical assessments.csv", 
    col_types = cols(`Date of Sale` = col_date(format = "%m/%d/%Y"))) %>% 
  clean_names()
head(df)
str(df)
```

```{r}
#data cleaning
#remove "C31" Replace with " "
df$outstanding_loan_balance <- str_replace(df$outstanding_loan_balance, 
                                           "C31", " ") %>% as.numeric()
#date
df$date_of_sale <- ymd(df$date_of_sale)
```

### Which region has the highest Outstanding Loan Balance exposure. Provide reasons and evidence of the analysis. Explain in approximately no less than (80 words)

```{r}
df %>% select(region, outstanding_loan_balance) %>% 
      group_by(region) %>% 
      summarise(Loan_Balance=sum(outstanding_loan_balance)) %>% 
       arrange(desc(Loan_Balance))

# as seen in the calculation,Suwami is leading in the outstanding loan balance by 8674531 followed by,Bumasi	8455168	, then Nilmark	5496709	 and finnally	
# Bira	1002851
```

### 

### Giving reasons and rank the region that is highly affected by fraud.Show evidence and explain in approximately (70 words)

```{r warning=FALSE, message=FALSE}
df %>%  select(region,investiagtion_outcome) %>% 
  filter(investiagtion_outcome== "Confirmed to be Fraud") %>% 
       group_by(region) %>% 
      summarise(Fraud_occurrence=n()) %>% arrange(desc(Fraud_occurrence))

# df %>%  pull(investiagtion_outcome) %>% unique()
```

### 

### Using the data determine the most affected phone model, and in which region. Show evidence and explain in approximately (70 words).

```{r warning=FALSE, message=FALSE}
df %>% select(region, model,investiagtion_outcome ) %>% 
        filter(investiagtion_outcome=="Confirmed to be Fraud") %>% 
        group_by(region, model) %>% 
        summarise(Leading_model=n()) %>%  
            arrange(desc(Leading_model))
```

### Using the data show the most affected month by fraud. Show evidence and explain in approximately (50 words).

```{r}
df <- df %>% mutate(month= month(date_of_sale, label=T))# %>% colnames()
# head(df$month)
d <- df %>%  select(month, investiagtion_outcome) %>% 
  filter(investiagtion_outcome=="Confirmed to be Fraud") %>% 
 group_by(month) %>% 
   summarise(NO_Fraud_month=n()) %>% arrange(desc(NO_Fraud_month))

d
```

### How could we potentially improve the fraud identification process? explain in approximately (100 words).

### What operational improvements should we investigate to improve the fraud investigation process? explain in approximately (100 words).

### Write an SQL query to replicate the results in Data-Sheet but only getting results for Suwami reg. Use the data on sheet named "Short schema".

### Show the process you used to clean the data. Show evidence and explain in approximately (80 words)
